Supplementary material

### 2021-06-28 submission/supplement.rmd
work on gusto part

>>>>

# Case study

We fitted a logistic regression model to predict 30-day mortality after a
myocardial infarction using data from the GUSTO-I trial. The risk factors
modeled were age, Killip class, systolic blood pressure heart rate, primary MI
and MI location. We restricted systolic blood pressure values to 120 mmHg and
modeled heart rate using linear spline transformation with a knot at 50 beats
per minute. A brief overview of the distribution of baseline characteristics is
presented in Table \@ref(tab:table1). The coefficients of the model can be seen
in Table \@ref(tab:modelCoefficients). We used this model in the manuscript to
derive baseline risk predictions to all patients of the study. For all patients
we set the treatment indicator to 0, assuming a constant relative treatment
effect. The model displayed very good internal validity (Table
\@ref(tab:internalValidity)).

```{r table1, echo=FALSE, warning=FALSE, message=FALSE}
load(here::here("data/raw/gusto.rda"))
gusto <- gusto %>%
  tibble() %>%
  filter(!is.na(tpa))

gustoTable1 <- gusto %>%
  mutate(
    tpa = factor(
      tpa,
      levels = c(0, 1),
      labels = c(
        "Tissue plasminogen activator",
        "Streptokinase"
      )
    ),
    Age = age,
    Outcome = factor(
      day30,
      levels = c(0, 1),
      labels = c(
        "Alive",
        "Dead"
      )
    )
  )

res <- table1::table1(
  ~ sex + Age + Killip + sysbp + + miloc + pulse + pmi + Outcome | tpa,
  data = gustoTable1,
  caption = "Baseline characteristics of the patients included in GUSTO-I trial."
)

table1::t1kable(res)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Smooth calibration of the derived prediction model."}
gustoModelData <- gusto %>%
  rename(
    "outcome" = "day30",
    "treatment" = "tpa"
  )

prediction <- glm(
  outcome ~ treatment + age + Killip + pmin(sysbp, 120) + rms::lsp(pulse, 50) + pmi + miloc,
  data = gustoModelData,
  maxit = 99,
  family = "binomial"
)

predictionStats <- rms::val.prob(
  y = gustoModelData$outcome, 
  logit = prediction$linear.predictors,
  logistic.cal = FALSE,
  statloc = FALSE
)
```

```{r modelCoefficients, echo=FALSE, warning=FALSE, message=FALSE}

predictionSummary <- summary(prediction)
table <- predictionSummary$coefficients

rownames(table)[2] <- "Treatment"
rownames(table)[3] <- "Age"
rownames(table)[4] <- "KILLIP (II)"
rownames(table)[5] <- "KILLIP (III)"
rownames(table)[6] <- "KILLIP (IV)"
rownames(table)[7] <- "Systolic blood pressure"
rownames(table)[8] <- "Pulse rate"
rownames(table)[9] <- "Pulse rate'"
rownames(table)[10] <- "Prior MI (yes)"
rownames(table)[11] <- "MI location (other)"
rownames(table)[12] <- "MI location (anterior)"

 knitr::kable(
    table,
    format      = "latex",
    align       = "r",
    booktabs    = TRUE,
    digits      = 4,
    format.args = list(
      big.mar      = ",",
      decimal.mark = "."
    ),
    caption    = "Coefficients of the prediction model for 30-day mortality after myocardial infarction (MI)."
  )

```

```{r internalValidity, echo=FALSE, warning=FALSE, message=FALSE}
table <- tibble(
  Metric = c(
    "AUC",
    "Brier score",
    "Emax",
    "E90",
    "Eavg"
  ),
  Value = c(
    predictionStats[2],
    predictionStats["Brier"],
    predictionStats["Emax"],
    predictionStats["E90"],
    predictionStats["Eavg"]
  )
)

 knitr::kable(
    table,
    format      = "latex",
    align       = "r",
    booktabs    = TRUE,
    digits      = 4,
    format.args = list(
      big.mar      = ",",
      decimal.mark = "."
    ),
    caption    = "Evaluation statistics of the prediction model for 30-day mortality in GUSTO-I."
  ) 
```
